Testing the Relationship between Phenomenological Control related to Illusion Sensitivity

### 2025-03-24 paper/R1/manuscript.qmd
minor changes to manuscript formatting

Interestingly, the fact that inter-individual variability in illusion sensitivity seems to persist in this task suggests that it is not solely explained by metacognitive abilities difference, and gives rise to the following question: is the variability in illusion sensitivity related to low-level perceptual processes (e.g., baseline precision of perceptual priors), or rather to the ability to actively control and "resist" the illusion in a top-town fashion in order to achieve the task at hand (higher-level modulation of the perceptual inference parameters). If the latter is true, then illusion sensitivity measured in contexts with strong task-demand characteristics, e.g., in paradigms where participants' performance is explicitly or implicitly assessed (i.e., where there is an incentive to downplay the illusion effect) might correlate with one's ability to alter one's subjective experience following suggestions - a mechanism referred to as "phenomenological control".

@@

The PCS will contain several manipulation check indices to identify problematic participants. **The phenomenological control task consists of various auditory and visual exercises. In one such exercise, participants receive the following instruction: “Open your eyes. You will see only two balls on the screen…just two balls”. However, three differently coloured balls are actually displayed. If participants select the option "no balls were shown", it indicates they failed to pay attention to both the auditory instructions and the visual stimuli. In another exercise, participants are asked to press the spacebar six times. If they press it fewer than five times within the allotted time, it suggests a lack of attentiveness to the auditory instructions. Participants will be excluded if they fail at least one of these checks. Participants will be excluded if they fail at least one of the attention checks.**

>>>>

Interestingly, the fact that inter-individual variability in illusion sensitivity seems to persist in this task suggests that it is not solely explained by **metacognitive ability differences**, and gives rise to the following question: is the variability in illusion sensitivity related to low-level perceptual processes (e.g., baseline precision of perceptual priors), or rather to the ability to actively control and "resist" the illusion in a **top-down** fashion in order to achieve the task at hand (higher-level modulation of the perceptual inference parameters). 
If the latter is true, then illusion sensitivity measured in contexts with strong task-demand characteristics, e.g., in paradigms where participants' performance is explicitly or implicitly assessed (i.e., where there is an incentive to downplay the illusion effect) might correlate with one's ability to alter one's subjective experience following suggestions - a mechanism referred to as "phenomenological control".

@@

The PCS will contain several manipulation check indices to identify problematic participants. 
**The phenomenological control task consists of various auditory and visual exercises.**
**In one such exercise, participants receive the following instruction: “Open your eyes. You will see only two balls on the screen…just two balls”.**
**However, three differently coloured balls are actually displayed.**
**If participants select the option "no balls were shown", it indicates they failed to pay attention to both the auditory instructions and the visual stimuli.**
**In another exercise, participants are asked to press the spacebar six times.**
**If they press it fewer than five times within the allotted time, it suggests a lack of attentiveness to the auditory instructions.**
**Participants will be excluded if they fail at least one of these checks. Participants will be excluded if they fail at least one of the attention checks.**
